---
title: "scor_01"
author: "NN"
date: "2023-01-02"
output: html_document
---

```{r setup}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load("E:\\workdata\\707957\\ID-NN\\Results\\Data\\scor.Rdata")
library(tidyverse)
library(psych)
library(lattice)
library(LMMstar)
library(readxl)
library(gtsummary)
library(ggplot2)
library(knitr)
library(writexl)

#setwd("E:/workdata/707957/ID-NN/Results/Til hjemsendelse/Final")
```
Calculating average number of entries per individual
```{r}

n_entries_per_individual <- plyr::count(scor$pnr)
n_entries_per_individual <- rename(n_entries_per_individual, pnr = x)
scor <- left_join(scor, n_entries_per_individual, by="pnr", all=T)
scor <- rename(scor, n_entries_per_individual = freq)

rm(n_entries_per_individual)
````

## Smaller dataframe for working on this project
```{r}
df <- scor %>% 
  select(pnr,
         statusdato, 
         indberetningsaar,
         perio,
         gingivitis,
         dmfs_perm,
         dmfs_temp, dmfs,
         ageatexam,
         birth_year,
         visit_year,
         koen,
         n_entries_per_individual)


```

#Finding mean(sd) number of entries per individual by birth cohorts 

```{r}

#Creating temporary scor variable with only one entry per individual
temp <- df%>%
  group_by(pnr) %>%
  slice_head()

describeBy(temp$n_entries_per_individual)

## Calculating mean(sd) by birth cohort
temp <- describeBy(temp$n_entries_per_individual, group = temp$birth_year)
temp <- do.call("rbind", temp) %>% 
  tibble::rownames_to_column() %>% 
  rename(birth_year = rowname) %>% 
  filter(birth_year >=1965 & birth_year <= 2010)

temp$birth_year <- as.numeric(temp$birth_year)

##plotting it
  ggplot(temp, aes(x = birth_year,
                 y = mean)) +
      geom_errorbar(aes(ymin = mean - sd,
                        ymax = mean+sd))+
                      geom_point() +
    xlab("Year of birth")+ ylab("Mean (sd)")+
    theme_classic()+
    scale_x_continuous(breaks = seq(1965, 2015, 5))
    ggsave("mean_sd_entries_by_birth_year.png")
    
rm(temp)
```

## Calculating number of visits by age group
```{r}
options(scipen = 999)

visits_df <- df %>%
  count(pnr, ageatexam) %>%
  mutate(age_groups = cut(ageatexam, breaks = c(0, 6, 11, 14, 19))) %>%
  pivot_wider(names_from = age_groups,
              values_from = n,
              names_prefix = "visit_at_") %>%
  select(!c(visit_at_NA)) %>%
  group_by(pnr) %>%
  summarise(across(starts_with("visit_at_"), sum, na.rm = TRUE)) %>%
  ungroup 

visits_df[apply(visits_df, 1, function(row) all(row !=0)), ] #Removing all rows with zero in, left with a total♣of 1.321.561 rows

#This means that only 100/3.510.760*1.321.561 = % of all individuals have a full set of registrations

#How many has three or more registrations?
visits_df$`visit_at_(0,6]`[visits_df$`visit_at_(0,6]` > 0] <- 1
visits_df$`visit_at_(6,11]`[visits_df$`visit_at_(6,11]` > 0] <- 1
visits_df$`visit_at_(11,14]`[visits_df$`visit_at_(11,14]` > 0] <- 1
visits_df$`visit_at_(14,19]`[visits_df$`visit_at_(14,19]` > 0] <- 1

sum_visits <- visits_df %>% 
  count(`visit_at_(0,6]`,`visit_at_(6,11]`, `visit_at_(11,14]`, `visit_at_(14,19]`)

sum_visits <- sum_visits %>% 
  mutate(percent_n = n/sum(n)*100)  

#Getting proportion visits by birth year

temp <- left_join(temp, visits_df, by ="pnr")

temp %>% 
  select(birth_year,
         `visit_at_(0,6]`,
         `visit_at_(6,11]`,
         `visit_at_(11,14]`,
         `visit_at_(14,19]`) %>%
  tbl_summary(
    by = birth_year) %>% 
  add_overall()

rm(temp, visits_df)
```
###Creating variable with only one entry per individual per age (0 to 18)
#note less than 10% loss of information doing this (apx. 8% to be specific, 27.144.696 entries vs. 24.988.782)
```{r}

restricted.df <- df %>% 
  group_by(pnr, ageatexam) %>%
  slice_head()

```
## Calculating average number of entries per individual for restricted data
```{r}

n_entries_per_individual <- plyr::count(restricted.df$pnr)
n_entries_per_individual <- rename(n_entries_per_individual, pnr = x)
restricted.df <- left_join(restricted.df, n_entries_per_individual, by="pnr", all=T)
restricted.df <- rename(restricted.df, n_entries_per_individual = freq)


#Creating df with only one entry per individual
temp <- restricted.df %>%
  group_by(pnr) %>%
  slice_head()

## Calculating mean(sd) by birth cohort
describeBy(temp$n_entries_per_individual)

## Calculating mean(sd) by birth cohort
temp <- describeBy(temp$n_entries_per_individual, group = temp$birth_year)
temp <- do.call("rbind", temp) %>% 
  tibble::rownames_to_column() %>% 
  rename(birth_year = rowname) %>% 
  filter(birth_year >=1965 & birth_year <= 2010)

temp$birth_year <- as.numeric(temp$birth_year)

##plotting it
  ggplot(temp, aes(x = birth_year,
                 y = mean)) +
      geom_errorbar(aes(ymin = mean - sd,
                        ymax = mean+sd))+
                      geom_point() +
       xlab("Year of birth")+ ylab("Mean (sd)")+
       theme_classic()+
    scale_x_continuous(breaks = seq(1965, 2010, 5))
    ggsave("mean_sd_entries_by_birth_year_restricted.png")


```

## Entries by age and year of report/birth/visit using restricted.df, i.e. how many individuals have at least one visit at a given age.
```{r}
#Number of visits by year of report and age
ind.age <- data.frame(unclass(table(restricted.df$indberetningsaar, restricted.df$ageatexam)))
ind.age[ind.age <= 5] <- 0
write_xlsx(ind.age, "E:\\workdata\\707957\\ID-NN\\Results\\Til hjemsendelse\\regyear_ageatexam.xlsx")
    
#Number of visits by year of birth and age
birth.age <-data.frame(unclass(table(restricted.df$birth_year, restricted.df$ageatexam)))
birth.age[birth.age <= 5] <- 0
  write_xlsx(birth.age, "E:\\workdata\\707957\\ID-NN\\Results\\Til hjemsendelse\\birthyear_ageatexam.xlsx")
  
#Number of visits by year of visit and age
visit.age <-data.frame(unclass(table(restricted.df$visit_year, restricted.df$ageatexam)))
visit.age[visit.age <= 5] <- 0
  write_xlsx(visit.age, "E:\\workdata\\707957\\ID-NN\\Results\\Til hjemsendelse\\visityear_ageatexam.xlsx")

#Number of visits by age
  age.res <-data.frame(table(restricted.df$ageatexam)) %>% 
  write_xlsx("E:\\workdata\\707957\\ID-NN\\Results\\Til hjemsendelse\\age_res.xlsx")
  
#Number of individuals with a visit by age
  age_total <- data.frame(table(df$ageatexam)) %>% 
  write_xlsx("E:\\workdata\\707957\\ID-NN\\Results\\Til hjemsendelse\\age_total.xlsx")

  rm(birth.age, ind.age, visit.age)
```  
  ##Plotting number of visits per year of report by age (actual enrollment into SCOR)
```{r}
 ##Plotting number of visits per year of report by age
ind.age <- data.frame(table(restricted.df$indberetningsaar, restricted.df$ageatexam)) %>% 
  rename(., Age = Var2)
ind.age[ind.age <= 5] <- 0
ind.age$Var1 <- as.character(ind.age$Var1)
ind.age$Var1 <- as.numeric(ind.age$Var1)

#Plotting it
ggplot(ind.age, aes(fill=Age, y = Freq, x = Var1))+
  geom_bar(position = "stack", stat = "identity")+
   xlab("Year of report")+ ylab("Frequency")+
      theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2021, 5))
ggsave("registrations_actual.png")

rm(ind.age)
```
```
##Plots of registrations by
```{r}
options(scipen = 999)

#Entries per individual total and restricted dataset:
age.res$Dataset[age.res$Freq > 0] <- "Restricted"
age_total$Dataset[age_total$Freq > 0] <- "Total"
age.df <- rbind(age.res, age_total) 
age.df$Dataset <- as.factor(age.df$Dataset)

#plotting it
age_full_res <- ggplot()+
    geom_line(data = age.df, aes(x=Var1, y=Freq, group = Dataset, color = Dataset))+
     xlab("Age")+ ylab("Number of visits")+
    theme_classic()
ggsave("Age.png")


#By Age before/after 1993

year_of_report <- data.frame(table(restricted.df$indberetningsaar, restricted.df$ageatexam))

year_of_report$Var1 <- as.character(year_of_report$Var1)
year_of_report$Var1 <- as.numeric(year_of_report$Var1)

year_of_report$Var2 <- as.character(year_of_report$Var2)
year_of_report$Var2 <- as.numeric(year_of_report$Var2)

year_of_report_pre <- year_of_report %>% 
  filter(Var1 < 1993) %>% 
  group_by(Var2) %>% 
  mutate(., visits = sum(Freq)) %>% 
  slice_head() %>% 
  mutate(., percent =(100/14029961*visits)) %>% 
  mutate(., group = "Before 1993")

year_of_report_post <- year_of_report %>% 
  filter(Var1 >= 1993) %>% 
  group_by(Var2) %>% 
  mutate(., visits = sum(Freq)) %>%
  slice_head() %>% 
  mutate(., percent =(100/14029961*visits)) %>% 
  mutate(., group = "After 1993")

pre_post <- rbind(year_of_report_pre, year_of_report_post)
pre_post$group <- as.factor(pre_post$group)

age_1993 <- ggplot()+
  geom_line(data = pre_post, aes(Var2, percent, group = group, color = group))+
  labs(x = "Age", y = "Percent of visits", color = "Period")+
 theme_classic()+
    scale_x_continuous(breaks = seq(0, 18, 1))
ggsave("age_1993.png")

#Bind with restricted/full data entries

setwd("E:/workdata/707957/ID-NN/Results/Scripts")

ggarrange(age_full_res, age_1993)
ggsave("age_plots.png")

rm(year_of_report, year_of_report_pre, year_of_report_post, age_1993, age.df, age_total, age.res)

````
#Plot visits by
```{r}

#Birth year
birth_year <- data.frame(table(restricted.df$birth_year))
birth_year[birth_year <= 5] <- 0

birth_year$Var1 <- as.character(birth_year$Var1)
birth_year$Var1 <- as.numeric(birth_year$Var1)

ggplot(data = birth_year, aes(x=Var1, y=Freq))+
    geom_bar(stat = "identity")+
  xlab("Year of birth")+ ylab("Frequency")+
    theme_classic()+
    scale_x_continuous(breaks = seq(1954, 2010, 5))
ggsave("birth_year.png")

#Year of report
year_of_report <- data.frame(table(restricted.df$indberetningsaar))

year_of_report$Var1 <- as.character(year_of_report$Var1)
year_of_report$Var1 <- as.numeric(year_of_report$Var1)

ggplot(data = year_of_report, aes(x=Var1, y=Freq, group = 1))+
    geom_line()+
   xlab("Year of report")+ ylab("Frequency")+
      theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2021, 5))
ggsave("year_of_report.png")

#Year of visit

year_of_visit <- data.frame(table(restricted.df$visit_year))

year_of_visit$Var1 <- as.character(year_of_visit$Var1)
year_of_visit$Var1 <- as.numeric(year_of_visit$Var1)

ggplot(data = year_of_visit, aes(x=Var1, y=Freq, group = 1))+
    geom_line()+
   xlab("Year of visit")+ ylab("Frequency")+
      theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2021, 5))
ggsave("year_of_visit.png")

rm(year_of_report, birth_year, year_of_visit, age.df, age, age_total)
```
##Plot gingivitis incidence rate by sex over time
```{r}
##Coding variable with first entry per individual with gingivitis otherwise just first entry.
df1 <- df %>% 
  group_by(pnr) %>% 
  slice(which.max(gingivitis))

#Making incidence plots
tmp_overall <- df1 %>% 
    group_by(indberetningsaar) %>% 
    summarise(cases_inc = sum(gingivitis),
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000*cases_inc/population) %>% 
    filter(cases_inc >=5 & indberetningsaar <2010)

tmp_sex <- df1 %>% 
    group_by(indberetningsaar, koen) %>% 
    summarise(cases_inc = sum(gingivitis), 
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000 * cases_inc / population) %>% 
    filter(cases_inc >=5 & indberetningsaar <2010)

gingi.inci <- ggplot()+
    geom_line(data = tmp_overall, aes(indberetningsaar, Incidens)) +
    geom_line(data = tmp_sex, aes(indberetningsaar, Incidens, color = koen))+
     labs(x = "Year of report", y = "Incidence", color = "Sex")+
     theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2010, 5))
ggsave("Gingi_inci.png")
```
#Gingi prevalence by age
```{r}

gingi_filter <- restricted.df %>% 
  filter(ageatexam == 5 | ageatexam == 9 | ageatexam == 15)

gingi_filter$ageatexam <- as.factor(gingi_filter$ageatexam)

#Making prevalence plot

tmp_age <- gingi_filter %>% 
    group_by(indberetningsaar, ageatexam) %>% 
    summarise(cases_inc = sum(gingivitis), 
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000 * cases_inc / population) %>% 
    filter(indberetningsaar <2010)


gingi.prev <- ggplot(data = tmp_age, aes(indberetningsaar, Incidens, color = ageatexam))+
    geom_line()+
     labs(x = "Year of report", y = "Prevalence", color = "Age")+
     theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2010, 5))
ggsave("Gingi_prev.png")

rm(gingi_filter)
```
#Plotting perio incidence over time
```{r}
##Coding variable with first entry per individual with gingivitis otherwise just first entry.
df2 <- df %>% 
  group_by(pnr) %>% 
  slice(which.max(perio))

tmp_overall <- df2 %>% 
    group_by(indberetningsaar) %>% 
    summarise(cases_inc = sum(perio),
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000*cases_inc/population) %>% 
    filter(cases_inc >=5 & indberetningsaar <2010)

tmp_sex <- df2 %>% 
    group_by(indberetningsaar, koen) %>% 
    summarise(cases_inc = sum(perio), 
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000 * cases_inc / population) %>% 
    filter(cases_inc >=5 & indberetningsaar <2010)

inci.perio <- ggplot()+
    geom_line(data = tmp_overall, aes(indberetningsaar, Incidens)) +
    geom_line(data = tmp_sex, aes(indberetningsaar, Incidens, color = koen))+
    labs(x = "Year of report", y = "Incidence", color = "Sex")+
    theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2010, 5))
ggsave("Perio_inci.png")
```
#Perio prevalence by age
```{r}

perio_filter <- restricted.df %>% 
  filter(ageatexam == 5 | ageatexam == 9 | ageatexam == 15)

perio_filter$ageatexam <- as.factor(perio_filter$ageatexam)

#Making prevalence plot

tmp_age <- perio_filter %>% 
    group_by(indberetningsaar, ageatexam) %>% 
    summarise(cases_inc = sum(perio), 
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000 * cases_inc / population) %>% 
    filter(indberetningsaar <2010)


prev.perio <- ggplot(data = tmp_age, aes(indberetningsaar, Incidens, color = ageatexam))+
    geom_line()+
     labs(x = "Year of report", y = "Prevalence", color = "Age")+
     theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2010, 5))
ggsave("Perio_prev_age.png")

rm(perio_filter)
```
# By sex: Prevalence plot dmf_s/DMF_S (Summed together) - number of individuals with 1 or more surfaces registered as destroyed, misssing or filled.
```{r}
#Coding binary dmfs variable to allow plotting
restricted.df$dmfs_bi <- ifelse(restricted.df$dmfs > 0, 1, 0)

##Coding variable with first entry per individual with gingivitis otherwise just first entry.
df3 <- restricted.df %>% 
  group_by(pnr) %>% 
  slice(which.max(dmfs_bi))

tmp_overall <- df3 %>% 
    group_by(indberetningsaar) %>% 
    summarise(cases_inc = sum(dmfs_bi),
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000*cases_inc/population) %>% 
    filter(cases_inc >=5 & indberetningsaar <2010)

tmp_sex <- df3 %>% 
    group_by(indberetningsaar, koen) %>% 
    summarise(cases_inc = sum(dmfs_bi), 
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000 * cases_inc / population) %>% 
    filter(cases_inc >=5 & indberetningsaar <2010)

inci.caries <- ggplot()+
    geom_line(data = tmp_overall, aes(indberetningsaar, Incidens)) +
    geom_line(data = tmp_sex, aes(indberetningsaar, Incidens, color = koen))+
    labs(x = "Year of report", y = "Incidence", color = "Sex")+
    theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2010, 5))
ggsave("dmfs_inci.png")

```

# By age: Prevalence plot dmf_s/DMF_S (Summed together) - number of individuals with 1 or more surfaces registered as destroyed, misssing or filled at age 5, 9 and 15.
```{r}

dmfs_filter <- filter(restricted.df, ageatexam == 5 | ageatexam == 9 | ageatexam == 15) #Filtering dataframe for only ages of interest

tmp_age <- dmfs_filter %>% 
    group_by(indberetningsaar, ageatexam) %>% 
    summarise(cases_inc = sum(dmfs_bi), 
              population = sum(n_distinct(pnr))) %>% 
    mutate(Incidens = 1000 * cases_inc / population) %>% 
    filter(cases_inc >=5 & indberetningsaar <2010)

tmp_age$ageatexam <- as.factor(tmp_age$ageatexam)

prev.caries <- ggplot(data = tmp_age, aes(x=indberetningsaar, y=Incidens, color = ageatexam))+
    geom_line()+
   labs(x = "Year or report", y = "Prevalence", color = "Age")+
      theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2021, 5))
ggsave("dmfs_prev.png")

rm(dmfs_filter)

```
#Binding inci and prev plots in one image
```{r}
#Binding all together into one image

ggarrange(gingi.inci, inci.perio, inci.caries,
          ncol = 1, nrow = 3)
ggsave("inci.png")


ggarrange(gingi.prev, prev.perio, prev.caries, 
          ncol = 1, nrow = 3)
ggsave("prev.png")

rm((inci.gingi, prev.caries, inci.perio, prev.perio, inci.caries, prev.caries)
```

#Comparison individuals DST and SCOR by year of birth 

```{r}

#Download DST excel file
#dst_births <- read_excel("C:/Users/FNOP707957.DSTFSE.010/Downloads/2022527121813375678117BEV3A.xlsx")

#Clean/format DST file
dst_births <- select(dst_births, -'Levendefødte og døde efter bevægelse og tid')
dst_births <- dst_births[-c(1,2,4,5),] 
rownames(dst_births) = c("n_births_dst")
colnames(dst_births) = c(
  "1954","1955","1956","1957","1958",
  "1959","1960","1961","1962","1963",
  "1964","1965","1966","1967","1968",
  "1969","1970","1971","1972","1973", 
  "1974","1975","1976","1977","1978",
  "1979","1980","1981","1982","1983",
  "1984","1985","1986","1987","1988",
  "1989","1990","1991","1992","1993",
  "1994","1995","1996","1997","1998",
  "1999","2000","2001","2002","2003",
  "2004","2005","2006","2007","2008",
  "2009","2010","2011","2012","2013",
  "2014","2015","2016","2017","2018",
  "2019","2020"
  )
dstbirths <- as.data.frame(t(dst_births))
dstbirths$n_births_dst <- as.numeric(dstbirths$n_births_dst)

#Merge with data on individuals per birth year from SCOR
SCOR_yearly_births <- data.frame(unclass(table(restricted.df$birth_year)))
SCOR_yearly_births[SCOR_yearly_births <= 5] <- 0
write_xlsx(SCOR_yearly_births, "E:\\workdata\\707957\\ID-NN\\Results\\Til hjemsendelse\\yearly births.xlsx")

n_births_comparison <- 
  merge(dstbirths, SCOR_yearly_births, by=0, all=TRUE)

colnames(n_births_comparison) = c("Year", "n_births_dst", "n_births_SCOR")

n_births_comparison <- n_births_comparison%>%
  na.omit(n_births_comparison)%>%
  mutate(difference = n_births_dst - n_births_SCOR)

#Trying without pre-1965 generations
n_births_comparison <- mutate(n_births_comparison[-c(1:11),])  
  
sum(n_births_comparison$n_births_SCOR)

sum(n_births_comparison$difference) # = minus62998

#Calculating dif in percent
100/sum(n_births_comparison$n_births_dst)*sum(n_births_comparison$difference)
```
## Calculating mean/SD entries for presence/abscence periodontitis and gingivitis, and min/max dmfs (based on highest achieved sum(DMF_S+dmf_s at any point in time during eligiblity for SCOR))
```{r}

#Summarising for average entries by gingivitis
summary <- summarize(n_entries_per_individual~birth_year+gingivitis, data=df1)

summary$gingivitis <- as.factor(summary$gingivitis)
summary$birth_year <- as.numeric(summary$birth_year)


summary <- summary %>% 
  filter(birth_year >= 1965 & birth_year <=2010)

#Plotting it
gingi <- ggplot()+
  geom_line(data = summary, aes(birth_year, mean, group = gingivitis, color = gingivitis))+
  labs(x = "Year of birth", y = "Mean entries", color = "Gingivitis")+
 theme_classic()+
    scale_x_continuous(breaks = seq(1965, 2010, 5))
ggsave("gingi_ulighed.png")


## For periodontitis coding variable with first entry per individual with periodontitis otherwise just first entry.
df2 <- df %>% 
  group_by(pnr) %>% 
  slice(which.max(perio))

#Summarising for average entries by periodontitis
summary <- summarize(n_entries_per_individual~birth_year+perio, data=df2)

summary$perio <- as.factor(summary$perio)
summary$birth_year <- as.numeric(summary$birth_year)


summary <- summary %>% 
  filter(birth_year >= 1965 & birth_year <=2010)

#Plotting it
perio <- ggplot()+
  geom_line(data = summary, aes(birth_year, mean, group = perio, color = perio))+
  labs(x = "Year of birth", y = "Mean entries", color = "Periodontitis")+
 theme_classic()+
    scale_x_continuous(breaks = seq(1965, 2010, 5))
ggsave("perio_ulighed.png")

#Creating df with only one entry per individual by highest dmfs_score
df3<- df%>%
  group_by(pnr) %>%
  slice(which.max(dmfs)) %>% 
  ungroup

df3$caries[df3$dmfs == 0] <- 0
df3$caries[df3$dmfs >=1 & df3$dmfs<=20] <- 1
df3$caries[df3$dmfs >=21 & df3$dmfs <=40] <- 2
df3$caries[df3$dmfs >=41 & df3$dmfs  >=60] <- 3
df3$caries[df3$dmfs >=61 & df3$dmfs >=80] <- 4
df3$caries[df3$dmfs >=81] <- 5


#Summarising for average entries by caries level

summary <- summarize(n_entries_per_individual~birth_year+caries, data=df3)

summary$caries <- as.factor(summary$caries)
summary$birth_year <- as.numeric(summary$birth_year)

summary <- summary %>% 
  filter(birth_year >= 1965 & birth_year <=2010)


#Plotting it

caries <- ggplot()+
  geom_line(data = summary, aes(birth_year, mean, group = caries, color = caries))+
    labs(x = "Year of birth", y = "Mean entries", color = "Caries level")+
 theme_classic()+
    scale_x_continuous(breaks = seq(1965, 2010, 5))
ggsave("caries_ulighed.png")


#Binding all together into one image

ggarrange(gingi, perio, caries,
          ncol = 1, nrow = 3)
ggsave("ulighed.png")

rm(summary)

```
#T-test of differences between groups, assume no difference between groups (two-sided test, 95 confint, equal variances)
```{r}
#Gingivitis
var.test(n_entries_per_individual~gingivitis, df1) #no sig. dif between variances
t.test(n_entries_per_individual~gingivitis, data = df1, var.eq = T)
#t = -205.7, df = 2968329, p-value < 0.00000000000000022, CI -1.270074 -1.246098, mean 8.968864/10.226950 

#Cohen d statistic of difference between two means
cohen.d(n_entries_per_individual~gingivitis, data = df1)

 #                        lower effect upper
#n_entries_per_individual  0.34   0.34  0.35

#Periodontitis
var.test(n_entries_per_individual~perio, df2)  #no sig. dif between variances
t.test(n_entries_per_individual~perio, data = df2, var.eq = T)
#t = 0.75452, df = 2968329, p-value = 0.4505, CI  -0.05268767  0.11864479, mean 9.144840/9.111862 

#Cohen d statistic of difference between two means
cohen.d(n_entries_per_individual~perio, data = df2)

 #                        lower effect upper
#n_entries_per_individual -0.03  -0.01  0.01


#Caries, using caries-free (group 0) as baseline for comparison

df01 <- df3 %>% 
  filter(caries == 0 | caries == 1)
t.test(n_entries_per_individual~caries, data = df01)
#t = -167.9, df = 687307, p-value < 0.00000000000000022,  CI -1.057141 -1.032746, mean 8.336859/9.381802 

#Cohen d statistic of difference between two means
cohen.d(n_entries_per_individual~caries, data = df01)

#                         lower effect upper
#n_entries_per_individual  0.28   0.28  0.28

df02 <- df3 %>% 
  filter(caries == 0 | caries == 2)
t.test(n_entries_per_individual~caries, data = df02)
#t = -83.006, df = 882748, p-value < 0.00000000000000022, CI -0.6376309 -0.6082136, mean 8.336859/8.959781 

cohen.d(n_entries_per_individual~caries, data = df02)

#                        lower effect upper
#n_entries_per_individual  0.17   0.17  0.18

df03 <- df3 %>% 
  filter(caries == 0 | caries == 3)
t.test(n_entries_per_individual~caries, data = df03)
#t = -13.135, df = 5032, p-value < 0.00000000000000022. CI  -0.7261924 -0.5375773, mean  8.336859/8.968744 

cohen.d(n_entries_per_individual~caries, data = df03)

#                         lower effect upper
#n_entries_per_individual  0.13   0.16  0.19


df04 <- df3 %>% 
  filter(caries == 0 | caries == 4)
t.test(n_entries_per_individual~caries, data = df04)
#t = -2.8887, df = 118.07, p-value = 0.004603, CI  -1.5709859 -0.2931117, mean 8.336859/9.268908 

cohen.d(n_entries_per_individual~caries, data = df04)

#                         lower effect upper
#n_entries_per_individual  0.06   0.24  0.42


df05 <- df3 %>% 
  filter(caries == 0 | caries == 5)
t.test(n_entries_per_individual~caries, data = df05)
#t = -8.6478, df = 923.22, p-value < 0.00000000000000022, CI -1.2537325 -0.7899413, mean  8.336859/9.358696

cohen.d(n_entries_per_individual~caries, data = df05)

#                         lower effect upper
#n_entries_per_individual   0.2   0.26  0.33
```
```{r}
#Entries per individual total and restricted dataset:
age.res$Dataset[age.res$Freq > 0] <- "Restricted"
age_total$Dataset[age_total$Freq > 0] <- "Total"
age.df <- rbind(age.res, age_total) 
age.df$Dataset <- as.factor(age.df$Dataset)

#plotting it
age_full_res <- ggplot()+
    geom_line(data = age.df, aes(x=Var1, y=Freq, group = Dataset, color = Dataset))+
     xlab("Age")+ ylab("Frequency")+
    theme_classic()
ggsave("Age.png")
```
## Expected visits

```{r}
#Making DF for expected enrollment

expected <- data.frame(Year=c(1972, 1973, 1974, 1975, 1976, 1977, 1978, 1979, 1980, 1981, 1982, 1983, 1984, 1985, 1986, 1987, 1988, 1989, 1990, 1991, 1992, 1993, 1994, 1995, 1996, 1997, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008),
X1=c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0 ,0 , 0, 0, 0, 0, 0, 0, 0, 0, 0),
X2=c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0 ,0 , 0, 0, 0, 0, 0, 0, 0, 0, 0),
X3=c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0 ,0 , 0, 0, 0, 0, 0, 0, 0, 0, 0),
X4=c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0 ,0 , 0, 0, 0, 0, 0, 0, 0, 0, 0),
X5=c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
X6 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X7=c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
X8=c(0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X9=c(0,0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X10=c(0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X11=c(0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X12=c(0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
X13=c(0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X14=c(0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X15=c(0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
X16=c(0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X17=c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
X18=c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))

colnames(expected) <- gsub("X","", colnames(expected))

exp <- pivot_longer(
expected,
!Year,
names_to = "Age", 
values_to = "Freq")

##Plotting it!
ggplot(expected, aes(fill=Age, y = Freq, x = Year))+
  geom_bar(position = "stack", stat = "identity")+
   xlab("Year of report")+ ylab("Frequency")+
      theme_classic()+
    scale_x_continuous(breaks = seq(1972, 2008, 5))
ggsave("enrollment_expected.png")

rm(expected)
```
